A comparison of classification algorithms applied to the Adult Database

### 2009-12-10 report/final.tex
correccoes + raw data no c4.5

This tool was used to apply the following algorithms: C4.5, ADTree, Random Forest, CART and Bayesian Networks.
Instead of using the command line tools, we used the Java API to build classifiers.

To use the data set we fetch it directly from the database and then apply various data pre-processing
techniques that are available within Weka:

@@

Performance metrics for both classification results will be registered: error rate and time spent
building the classifier and classifying all the examples.

@@

We executed the algorithm using different combinations of the parameters described above.
Table \ref{tbl:results_c45} displays the obtained results. From the table
we note that once the confidence threshold reaches a certain level ($\approx$0.55) the results
are very similar for the unpruned classifier and the over-fitting effect starts to get worse,
where the train data gets very good results ($\approx$11.19 error rate) and the test set gets a
large error rate ($\approx$24).

The confidence threshold for best test error rates appears to be in the interval 0.05-0.15.

\begin{table}[ht]
  \begin{center}
  \begin{tabular}{ | l | c | c | c | c | c |}
    \hline
    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} & \textbf{Nodes} & \textbf{Leaves} \\ \hline
    -U & 11.18 & 23.74 & 0m24.378s & 6919 & 6198 \\ \hline
    -C 0.05 & 15.67 & 22.86 & 0m24.414s & 46 & 38 \\ \hline
    -C 0.1 & 15.22 & 21.69 & 0m25.417s & 106 & 89 \\ \hline
    -C 0.15 & 14.85 & 22.25 & 0m25.086s & 186 & 158 \\ \hline
    -C 0.2 & 14.52 & 22.05 & 0m24.436s & 298 & 254 \\ \hline
    -C 0.25 & 14.40 & 22.08 & 0m24.628s & 376 & 319 \\ \hline
    -C 0.3 & 14.17 & 23.26 & 0m24.765s & 494 & 423 \\ \hline
    -C 0.35 & 13.84 & 22.42 & 0m25.154s & 692 & 593 \\ \hline
    -C 0.4 & 13.68 & 22.49 & 0m24.325s & 835 & 721 \\ \hline
    -C 0.45 & 13.26 & 22.43 & 0m25.086s & 1143 & 992 \\ \hline
    -C 0.5 & 12.95 & 22.51 & 0m24.874s & 1483 & 1292 \\ \hline
    -C 0.55 & 11.19 & 24.94 & 0m25.899s & 6727 & 6036 \\ \hline
    -C 0.6 & 11.19 & 23.94 & 0m26.882s & 6727 & 6036 \\ \hline
    -C 0.7 & 11.19 & 23.94 & 0m26.638s & 6727 & 6036 \\ \hline
    -C 0.8 & 11.19 & 23.94 & 0m26.818s & 6727 & 6036 \\ \hline
    -C 0.9 & 11.19 & 23.94 & 0m26.634s & 6727 & 6036 \\ \hline
  \end{tabular}
  \caption{Results for C4.5.}
  \label{tbl:results_c45}
  \end{center}
\end{table}

A decision tree built by this algorithm with the confidence threshold set to 0.05 is shown in Listing \ref{decision_tree_c45}.

>>>>

This tool was used to apply the following algorithms: C4.5, ADTree, Random Forest, CART and Bayesian Networks.
Instead of using the command line tools, we used the Java API to build classifiers.
We will build classifiers using only nominal attributes, with numeric values being
discretized into categories, and classifiers with the data unchanged, to test
the effects of using numeric attributes within the algorithms.

To use the data set we fetch it directly from the database and then optionally
apply various data pre-processing techniques that are available within Weka:

@@

Performance metrics for both classification results will be registered: error rate and time spent
building the classifier and classifying all the examples.
We will separate the results for the discretized data set and for the raw data set.

@@

We executed the algorithm using different combinations of the parameters described above.
Table \ref{tbl:results_c45_discr} and \ref{tbl:results_c45_raw} displays the results
with discretized and raw data, respectively. From the tables we can see that
using numeric attributes increases the number of nodes and leaves, and
that once the confidence threshold increases, the test error rate
increases while the train error rate decreases, which indicates over-fitting.
This effect is clearly visible for unpruned trees.

\begin{table}[ht]
  \begin{center}
  \begin{tabular}{ | l | c | c | c | c | c |}
    \hline
    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} & \textbf{Nodes} & \textbf{Leaves} \\ \hline
    -U & 11.18 & 23.74 & 0m24.378s & 6919 & 6198 \\ \hline
    -C 0.05 & 15.67 & 22.86 & 0m24.414s & 46 & 38 \\ \hline
    -C 0.1 & 15.22 & 21.69 & 0m25.417s & 106 & 89 \\ \hline
    -C 0.15 & 14.85 & 22.25 & 0m25.086s & 186 & 158 \\ \hline
    -C 0.2 & 14.52 & 22.05 & 0m24.436s & 298 & 254 \\ \hline
    -C 0.25 & 14.40 & 22.08 & 0m24.628s & 376 & 319 \\ \hline
    -C 0.3 & 14.17 & 23.26 & 0m24.765s & 494 & 423 \\ \hline
    -C 0.35 & 13.84 & 22.42 & 0m25.154s & 692 & 593 \\ \hline
    -C 0.4 & 13.68 & 22.49 & 0m24.325s & 835 & 721 \\ \hline
    -C 0.5 & 12.95 & 22.51 & 0m24.874s & 1483 & 1292 \\ \hline
  \end{tabular}
  \caption{Results for C4.5 with discretized attributes.}
  \label{tbl:results_c45_discr}
  \end{center}
\end{table}

\begin{table}[ht]
  \begin{center}
  \begin{tabular}{ | l | c | c | c | c | c |}
    \hline
    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} & \textbf{Nodes} & \textbf{Leaves} \\ \hline
    -U & 9.48 & 21.58 & 0m7.525s & 6422 & 5491 \\ \hline
    -C 0.05 & 13.78 & 19.49 & 0m7.667s & 149 & 105 \\ \hline
    -C 0.1 & 13.37 & 19.06 & 0m7.969s & 278 & 212 \\ \hline
    -C 0.15 & 12.99 & 18.84 & 0m7.708s & 361 & 278 \\ \hline
    -C 0.2 & 12.95 & 18.89 & 0m7.812s & 373 & 284 \\ \hline
    -C 0.25 & 12.86 & 18.86 & 0m7.974s & 433 & 331 \\ \hline
    -C 0.3 & 12.11 & 19.73 & 0m8.064s & 806 & 624 \\ \hline
    -C 0.35 & 12.02 & 19.77 & 0m8.035s & 878 & 678 \\ \hline
    -C 0.4 & 11.54 & 20.47 & 0m8.255s & 1264 & 995 \\ \hline
    -C 0.5 & 11.22 & 20.28 & 0m8.117s & 1603 & 1257 \\ \hline
  \end{tabular}
  \caption{Results for C4.5 with non-discretized attributes.}
  \label{tbl:results_c45_raw}
  \end{center}
\end{table}

A decision tree built by this algorithm with the confidence threshold set to 0.05 and
discretized data is shown in Listing \ref{decision_tree_c45}.
